### 2012-11-15 proposal/proposal.tex
Create initial LaTeX structure for proposal

% proposal.tex | CS 6260 - "Vendetta" Proposal

% setup
\documentclass[9pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{fullpage}
\usepackage{color}
\usepackage{enumerate}
\usepackage{hyperref}
\usepackage{mathtools}
\usepackage{listings}
\usepackage{array}
\usepackage{minted}
\usepackage{graphicx}
\usepackage{parskip}

% don't number sections
\setcounter{secnumdepth}{0}

% don't highlight links
\hypersetup{hidelinks}

% custom commands
\newcommand{\todo}[1]{\textcolor{red}{\textbf{TODO:} \emph{#1}}}
\newcommand{\term}[1]{\textit{#1}}
\newcommand{\bterm}[1]{\textbf{\textit{#1}}}
\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\super}[1]{\textsuperscript{#1}}
\newcommand{\sub}[1]{\textsubscript{#1}}

% title
\title{``Vendetta'': End-to-end auditable voting}
\date{}
\author{Kelsey Francis, Curtis Free, Christopher Martin}

% document
\begin{document}
	\maketitle

	\todo{Write the proposal.}
	\cite{dummy}

	\bibliography{sources}{}
	\bibliographystyle{plain}
\end{document}

% vim: set colorcolumn=101 cursorcolumn cursorline: